Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept

### 2021-10-31 paper/article.tex
Paper progress

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\begin{document}

@@

TODO: arch figure here (from Slide 5)

\subsection{Architecture Overview}
First, we present the architecture of our co-processor design. The co-processor is
composed of two artifical neural networks:
\begin{itemize}
	\item A stimulation and neural dynamics model, known as an ``Emulator Network''.  It models the relationship
\end{itemize}
* EN architecture
** Single layer LSTM with tanh activations, following by linear readout
** LSTMs proved to be necessary for tracking long-running neural dynamics in a way that gave us a useful training signal
*** Explore why? Revisit
** Inputs: brain observations, stimulation vector, trial end indicator
** Outputs: predicted muscle velocity
* CPN architecture
** Single layer LSTM with tanh activations, following by linear readout
** Inputs: brain observations, trial end indicator
** Outputs: stimulation vector

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\usepackage{graphicx}
\graphicspath{ {./figs/} }
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\begin{document}

@@

\begin{figure}
\includegraphics[width=\textwidth]{weill_arch.png}
\caption{Using a co-processor to drive external task performance after a traumatic brain injury}
\centering
\label{fig:weill}
\end{figure}

\subsection{Architecture Overview}
First, we present the architecture of our co-processor design. This design aims to solve two
fundamental challenges in using neural stimulation to improve external task performance.
First, neural networks exhibit long-running and highly nonlinear dependencies, necessitating
the need for a stimulation agent to account for-far distant effects of the stimulation it
applies. Second, with neurologically complex tasks, such as grasping, the mapping between
neural activity as-measured and the external task cannot be determined \textit{a priori}.

Our co-processor attempts to solve this with a pair of artificial neural networks:
\begin{itemize}
	\item A stimulation and neural dynamics model, known as an ``Emulator Network'' (EN). It models the relationship
	      between the stimulation, neural dynamics, and external task. It's purpose is for training the second network.
	\item A stimulation agent, known as the ``co-processor network'' (CPN), which maps neural activity, and possibly
	      data from external sensors, to stimulation parameters.
\end{itemize}

We co-train the EN and CPN, with the goal of training a CPN whose output stimulation parameters
improve task performance. By continually training both, they adapt to the brain as it changes,
effectively causing brain-stimulator co-adaptation. The EN is a tool for training the CPN,
giving us a way to back-propagate task error to the CPN. It outputs task-relevant
metrics - a prediction of muscle velocities in our case - given measurements of neural
activity, and the stimulation parameters output from the CPN. If the EN is trained in a
particular way, and to a sufficient level of precision, it can be used to overcome the
challenge of not knowing a mapping between stimulation and the task. When training the
CPN, we in-effect treat the EN's output as the true task performance, or a related
metric, and then backpropagate the loss defined in terms of that metric. See Figs X, Y.
We will illustrate the details of the training algorithm below.

In our present demonstration, the EN is constituted as a single layer, fully connected,
long short-term memory (LSTM) recurrent neural network (RNN), with hyperbolic tangent ($tanh$)
activations, and a linear readout. The general notion of a co-processor does
not require this precise architecture, but for our example, we found that the LSTM
approach allows the network to continuously adapt to long-running dependencies in
the simulated neural dynamics, far better than a vanilla RNN. The CPN is constituted
as an almost identical network, though with a different dimensionality, as we explain
below. There is no strict requirement for the EN and CPN to be so similar, but we
found this simple architecture to work well in our example.

Note that the EN effectively constitutes a stimulation model. Its design is
somewhat motivated by the common linear time-invariant state space model of
stimulation, as in e.g. Yang et al. \cite{shanechi.stimmodel}, which is a helpfully
simple (linear) approach. That approach seeks to model neural network dynamics in
terms of a stationary linear model, which, due in part to its simplicity and the
number of techniques developed for it, lends itself to useful interpretations and
analyses of those neural dynamics.

There are some key differences between that approach and the EN architecture we
present here. Our EN's purpose is simply to train the CPN, so we pick an
architecture we find to be best suited to that job. First, our EN architecture
is not linear, through the use of a $tanh$ activation function. That allows us
to escape some of the limits in representational power of linear models.
Second, LSTM cells are designed to better capture long-running network
dynamics, and in our case we found that they were crucial for that reason.
Compare that to the simpler cells commonly used in vanilla RNNs, which
yield functionally the same approach as the linear state space model, less
the use of a nonlinear activation function.

// TODO? Try having no nonlinearity?

@@

* Overview of the four experiments:
** Vanilla
** Co-adaptive with connectivity loss
** Co-adaptive with partial M1 loss
** Co-adaptive with data reuse
